Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-10-21 apj.tex
respond to Greg and Gerard's VC comments

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The two color parameters exhibit the behavior expected from the two-parameter
\citet{1999PASP..111...63F} dust model, yielding a per-supernova scatter in
total-to-selective extinction, an effective
%-----
$R^F_V \sim 2.4$
%------
for the full sample, and a difference in the average $R^F_V$ between extremely blue and red supernovae .  The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
the two parameters
plus a detection of a third parameter at  $>98$\% confidence; this new parameter is responsible for supernova color diversity
beyond those tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

Combining the effects of both parameters, we can define a total $X$-band extinction $A_X=\gamma^0_X k_0 + \gamma^1_X k_1$
and a corresponding $E(B-V) = A_B-A_V$.  Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of $R_V=2.77^{+0.08}_{-0.09}$.  The fit also gives the zeropoint, which we use to 
estimate $R_V=(A_V-zeropoint)/E(B-V)$ per supernova, shown in Figure~\ref{avebv:fig}.
Extremely blue supernovae ($E(B-V) <-0.05$ on the plot) have an average $R_V=1.88 \pm   0.13$,
whereas the extreme red objects ($E(B-V) >0.1$) have $R_V=  2.96 \pm   0.11$, indicating that different physical mechanisms
operate in these regimes.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/avebv.pdf}
   \includegraphics[width=2.8in]{output11/rv.pdf}
   \caption{Left: $A_V$ versus $E(B-V)$ and Right: $R_V=A_V/(E(B-V)-zeropoint$\
   versus $E(B-V)$ for the supernova sample.
   \label{avebv:fig}}
\end{figure}

Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
parameters of a single  dust model.
\color{red}
Plotted in Figure~\ref{avebv:fig} are the expectations for measured $A(V)$, $E(B-V)$, and $R_V$ for $R^F_V=1.9, 2.4, 3.1$
within the context of the \citetalias{1999PASP..111...63F} model.
Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
non-linear, and $R_V$ is not constant, for a fixed value of $R^F_V$.
\color{black}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/plane0.pdf}
   \includegraphics[width=2.8in]{output11/plane1.pdf}
   \caption{Two different perspectives in $UVI$-space of the vectors $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.4$ is shown in the dot-dashed
   line.
   \label{plane:fig}}
\end{figure}

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant changes of note are in the standard
deviations of the residuals in the $V$ and $I$ bands, which are now reduced.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust law, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.22 & 2.75 \\
0.80 & -5.51
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors $\vec{\epsilon}$ are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 3\% and 4\% to the total  length of the vector.
The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.

>>>>

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The two color parameters exhibit the behavior expected from the two-parameter
\citet{1999PASP..111...63F} dust model, yielding a per-supernova scatter in
total-to-selective extinction.  The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
the two parameters
plus a detection of a third parameter at  $>98$\% confidence; this new parameter is responsible for supernova color diversity
beyond those tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

\color{red}
Recall that our model does not specify the color of a supernova that suffers no extrinsic effects, but rather sets the
mean of $k_0$ and $k_1$ of the distribution to zero.
There is an unconstrained $k_0^0$, $k_1^0$ that corresponds to an intrinsic supernova.  The true extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true color excess is $E^T(B-V) = A^T_B-A^T_V$.
Despite this ambiguity, a linear slope relating the $E^T(B-V)$ and $A^T_V$ for a set of supernova is the same slope
relating $A_V$ and $E(B-V)$.
\color{black}
Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of $R_V=2.77^{+0.08}_{-0.09}$,
\color{red}
which is the observed broad-band total-to-selective extinction exhibited by our sample.
In contrast, the quantity $R^F_V = A^F_V/E^F(B-V)$ for an individual supernova is unspecified without knowledge of the intrinsic color.
For ta qualitative view, we choose $k_0$  and $k_1$ such that the bluest supernova roughly correspond to $E(B-V)=0$.
With this ansatz, we
estimate $R^T_V$ per supernova as shown in Figure~\ref{avebv:fig}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/avebv.pdf}
   \includegraphics[width=2.8in]{output11/rv.pdf}
   \caption{
	\color{red}
   Left: $A_V$ versus $E(B-V)$, and the slope $R_V$ of the linear fit to the data. Right: Shown for qualitative purposes,
   $R^T_V=A^T_V/E^T(B-V)$ 
   versus $E^T(B-V)$ for the supernova sample, given an ansatz for the definition of an intrinsic supernova.  Overplotted is the mean and standard deviation of those points in equally divided bins.
   On both plots are overplotted the expected values for the \citetalias{1999PASP..111...63F} model
   for specific choices of $R^F_V$, normalized to have $A_V=0$ at $E(B-V)=0$.  Note that these lines are not straight on the left plot, and are not constant on the plot to the right.
   \color{black}
   \label{avebv:fig}}
\end{figure}

Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
parameters of a single  dust model.
\color{red}
Plotted in Figure~\ref{avebv:fig} are the expectations for measured $A(V)$, $E(B-V)$, and $R^T_V$ for $R^F_V=1.1, 1.9, 2.5, 3.1$
within the context of the \citetalias{1999PASP..111...63F} model.
Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
non-linear, and $R^T_V$ is not constant, for a fixed value of $R^F_V$.
\color{black}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/plane0.pdf}
   \includegraphics[width=2.8in]{output11/plane1.pdf}
   \caption{
   \color{red}
   Two different perspectives in $UVI$-space of the vectors $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.4$ is shown in the dot-dashed
   line, almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.
   \color{black}
   \label{plane:fig}}
\end{figure}

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant changes of note are in the standard
deviations of the residuals in the $V$ and $I$ bands, which are now reduced.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust law, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.22 & 2.75 \\
0.80 & -5.51
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors $\vec{\epsilon}$ are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 3\% and 4\% to the total  length of the vector.
%The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.
